Cancer Applications of Choropleth Maps, and the Potential of Cartograms and Alternative Map Displays

### 2019-12-17 Response/Response.Rmd

**Editor suggestion:**
**We would suggest that the aim of the manuscript could be phrased to highlight the current approaches and limitations of typical visualisation methods for cancer maps, along with outlining possible alternative methods and challenges in implementing these?**

We have restructed our original submission to provide a clear path from current approaches used, to the alternative approaches that should be explored. We highlight the limitations of using current display methods, to encourage the use of alternative display methods. We also present the challenges that may be faced when implementing the alternative designs.

The manuscript now contains the following sections:

**Traditional approaches for cancer map displays:** This section discusses the choropleth map as the common map display used in cancer maps and atlases.
*Limitations of choropleth displays:* This section presents the possible issues with displaying cancer statistics on a choropleth map.
*Contemporary alternatives to choropleth maps:* We have shown that there are many alternative displays that highlight population dense regions.
*Challenges of alternative approaches:* Alternative displays also have limitations when displaying spatial distributions of cancer statistics.

@@

Bowel Cancer Australia Atlas:
...shows the average Standardised Incidence Ratio of colorectal cancer for Australian males from 2006 to 2010 in Australia. It is published by Bowel Cancer Australia https://www.bowelcanceraustralia.org/.

*Reviewer A: *
*Major*
*Comment 1: *
The Abstract should be more focused- perhaps give an example of how cartogram mapping has been applied to data visualisation. What are the benefits of this approach and suggestions for the future.

*Reply 1:* 
We agree with the reviewer. Alternative mapping displays have been implemented in political contexts. Choropleth maps have often been used.

Cancer atlases communicate cancer statistics over geographic domains. They subdivide these domains into administrative regions such as countries, states, or suburbs. When communicating human-related statistics in Australia, the geographic map base draws attention to sparsely populated rural areas. The smaller geographic areas may not draw the attention of readers, but they are important to consider if they are more densely populated. Alternative map displays can be more effective than a choropleth map at revealing spatial patterns. Alternative displays have rarely been used in the context of online cancer atlases, but the Globocan atlas, [@Globocan] provides an example of how alternative statistical displays can be implemented.

@@

Cancer atlases communicate cancer statistics over geographic domains. They subdivide these domains into administrative regions such as countries, states, or suburbs. When communicating human-related statistics in Australia, the geographic map base draws attention to sparsely populated rural areas. The smaller geographic areas may not draw the attention of readers, but they are important to consider if they are more densely populated. Alternative map displays can be more effective than a choropleth map at revealing spatial patterns. Alternative displays have rarely been used in the context of online cancer atlases, but the Globocan atlas, [@Globocan] provides an example of how alternative statistical displays can be implemented.

*Comment 3: *

3. Entire paper needs substantial English language editing

*Reply 3:* 
Thank you for this point. Subtantial time has been taken to improve the language used.

### 2019-12-17 review_article.Rmd
colour scales

Cancer atlases communicate cancer statistics over geographic domains. They subdivide these domains into administrative regions such as countries, states, or suburbs. When communicating human-related statistics in Australia, the geographic map base draws attention to sparsely populated rural areas. The smaller geographic areas may not draw the attention of readers, but they are important to consider if they are more densely populated. Alternative map displays can be more effective than a choropleth map at revealing spatial patterns. Alternative displays are effective for the communication of election data aggregated at the state or electorate level on online news sites.

@@

Epidemiologists and statisticians have developed the statistics used to communicate the burden of cancer over several decades. Table 2<!--\ref{tab:measures}--> summarizes the measures commonly presented in published cancer atlases. Mortality rates are commonly presented as relative rates of risk across the population and age-adjusted to correct for the higher prevalence of cancers in older populations.
Howe [@HEDP] describes Stock's development of the standardized mortality ratios through the 1930s.

@@

Map users can define the population further using the demographics of the individuals within each area. Demographics that can influence cancer rates include information regarding the age and sex distribution of the areas presented. Some digital atlases surveyed allow users to control the display to select subsets such as males, females or those aged over 65. Similarly, socioeconomic indicators, such as unemployment rates, poverty rates, remoteness, and education levels, can explain how the experience of cancer prevalence varies for various members of society. Few atlases provide this
level of detail. Introducing population and demographic information helps to interpret the rates in areas effectively, but there will still be uncertainty around the rates. To address this, a cancer atlases often communicates uncertainty about the value of a statistic. There are several potential sources of uncertainty: sampling error, errors arising from the disease reporting process (or data collection), and errors arising from the statistical modeling or simulation process. The most common measures used to represent uncertainty were credible or confidence intervals (CIs). Displaying the uncertainty associated with reported statistics is a vital feature of a cancer map, but it is difficult to display effectively. The map focuses on displaying the statistic and lacks additional space to represent the uncertainty. Providing an adjacent map or overlaying maps with symbols [@VSSDCUC] are two approaches.

@@

Choropleth maps colour each geographic unit to allow map users to measure the value of the statistic [@EI]. Map users contrast the colours in neighbouring areas to understand the spatial distribution. Map creators must choose an attractive and effective colour scheme, but is a difficult task [@CB]. The ColorBrewer system supports map creators in the decision making process, but this does not consider perceptual colour
specifications.

>>>>

Cancer atlases communicate cancer statistics over geographic domains. They subdivide these domains into administrative regions such as countries, states, or suburbs. When communicating human-related statistics in Australia, the geographic map base draws attention to sparsely populated rural areas. The smaller geographic areas may not draw the attention of readers, but they are important to consider if they are more densely populated. Alternative map displays can be more effective than a choropleth map at revealing spatial patterns. Alternative displays have rarely been used in the context of online cancer atlases, but the Globocan atlas, [@Globocan] provides an example of how alternative statistical displays can be implemented.

@@

Epidemiologists and statisticians have developed the statistics used to communicate the burden of cancer over several decades. Table 2<!--\ref{tab:measures}--> summarizes the measures commonly presented in published cancer atlases. Mortality rates are commonly presented as relative rates of risk across the population and age-adjusted to correct for the higher prevalence of cancers in older populations. Stock developed the standardized mortality ratios through the 1930s, this is decribed by Howe [@HEDP].

@@

Map users can define the population further using the demographics of the individuals within each area. Demographics that can influence cancer rates include information regarding the age and sex distribution of the areas presented. Some digital atlases surveyed allow users to control the display to select subsets such as males, females or those aged over 65. Similarly, socioeconomic indicators, such as unemployment rates, poverty rates, remoteness, and education levels, can explain how the experience of cancer prevalence varies for various members of society. Few atlases provide this level of detail.

Introducing population and demographic information helps to interpret the rates in areas effectively, but there will still be uncertainty around the rates. To address this, a cancer atlases often communicates uncertainty about the value of a statistic. There are several potential sources of uncertainty: sampling error, errors arising from the disease reporting process (or data collection), and errors arising from the statistical modeling or simulation process. The most common measures used to represent uncertainty were credible or confidence intervals (CIs). Displaying the uncertainty associated with reported statistics is a vital feature of a cancer map, but it is difficult to display effectively. The map focuses on displaying the statistic and lacks additional space to represent the uncertainty. Providing an adjacent map or overlaying maps with symbols [@VSSDCUC] are two approaches.

@@

Choropleth maps colour each geographic unit to allow map users to measure the value of the statistic [@EI]. Map users contrast the colours in neighbouring areas to understand the spatial distribution. Map creators must choose an attractive and effective colour scheme, but is a difficult task [@CB]. The ColorBrewer system supports map creators in the decision making process, but this does not consider perceptual colour specifications. When communicating information using colour, a map creator should use a scheme that has a linear color gradient, with perceptually uniform color spaces to convey equal steps in data align with equal steps in the colour scheme [@PUCS].
